### 2015-08-18 article/cooccurrence-noise-experiment.tex
separate files for article text samples, document frequency function

\texttt{the \textcolor{red}{CAT\_2} domestic \textcolor{red}{CAT\_3} was \textcolor{red}{CAT\_3} first \textcolor{red}{CAT\_2} classified as felis catus\newline the semiferal \textcolor{red}{CAT\_2} \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_3} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_3} is \textcolor{red}{CAT\_1} one whose ancestry is recorded by a \textcolor{red}{CAT\_1} \textcolor{red}{CAT\_1} fancier \textcolor{red}{CAT\_1} organization\newline a purebred \textcolor{red}{CAT\_1} is one whose ancestry contains only \textcolor{red}{CAT\_1} individuals of the same breed\newline the \textcolor{red}{CAT\_2} \textcolor{red}{CAT\_1} skull is unusual among mammals in having very large eye \textcolor{red}{CAT\_3} sockets\newline another unusual feature \textcolor{red}{CAT\_2} is that the \textcolor{red}{CAT\_2} \textcolor{red}{CAT\_3} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} is usually dominant \textcolor{red}{CAT\_1} over the others\newline ...\newline the domestic dog \textcolor{red}{CAT\_3} canis \textcolor{red}{CAT\_1} lupus familiaris is a domesticated canid which has been selectively bred\newline dogs perform many roles for \textcolor{red}{CAT\_3} people such as \textcolor{red}{CAT\_3} hunting herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the \textcolor{red}{CAT\_2} time at \textcolor{red}{CAT\_1} which female dogs will have \textcolor{red}{CAT\_2} their first \textcolor{red}{CAT\_3} estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction}