### 2015-08-18 article/word-frequency-experiment-text-void.tex

\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned \textcolor{red}{VOID\_3} by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in \textcolor{red}{VOID\_1} having very large eye sockets\newline \textcolor{red}{VOID\_2} another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others\newline ...\newline the domestic dog canis lupus familiaris is a domesticated canid which has been selectively bred\newline \textcolor{red}{VOID\_1} dogs perform many roles for people such as hunting \textcolor{red}{VOID\_1} herding and pulling loads\newline in domestic dogs sexual maturity begins to happen around age six to twelve months\newline this is the time at which female dogs will have their first estrous cycle\newline some dog breeds have acquired traits through selective breeding that interfere with reproduction\newline ...\newline rabbits are small mammals in the family leporidae\newline the male is called a buck and the female is a doe\newline the european rabbit has been introduced to many places around the world\newline more than half \textcolor{red}{VOID\_1} the worlds rabbit population resides in north america\newline rabbits are \textcolor{red}{VOID\_1} prey animals and are therefore constantly aware of their surroundings\newline rabbits are most active at dawn and dusk\newline as with other prey animals rabbits often sleep with their eyes open}