### 2015-09-03 article/word-frequency-experiment-text-void.tex

\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned \textcolor{red}{VOID\_3} by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in \textcolor{red}{VOID\_1} having very large eye sockets\newline \textcolor{red}{VOID\_2} another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others}