### 2015-09-10 article/word-frequency-experiment-text-void.tex

\texttt{the domestic cat was first classified as felis catus\newline the semiferal cat a mostly outdoor cat is not owned by any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline \textcolor{red}{VOID\_1} within groups one cat is usually dominant over the others}