### 2015-10-01 article/word-frequency-experiment-text-cat.tex

\texttt{the domestic \textcolor{red}{CAT\_2} was first classified as felis catus\newline the semiferal \textcolor{red}{CAT\_1} a mostly outdoor \textcolor{red}{CAT\_1} is not owned by any one individual\newline a pedigreed \textcolor{red}{CAT\_1} is one whose ancestry is recorded by a \textcolor{red}{CAT\_2} fancier organization\newline a purebred \textcolor{red}{CAT\_2} is one whose ancestry contains only individuals of the same breed\newline the \textcolor{red}{CAT\_4} skull is unusual among mammals in having very large eye sockets\newline another unusual feature is that the \textcolor{red}{CAT\_1} cannot produce taurine\newline within groups one \textcolor{red}{CAT\_1} is usually dominant over the others}