### 2015-10-01 article/word-frequency-experiment-text-void.tex

\texttt{\textcolor{red}{VOID\_1} the domestic cat was first classified as felis catus\newline the semiferal cat \textcolor{red}{VOID\_3} a mostly outdoor cat is not \textcolor{red}{VOID\_2} owned by \textcolor{red}{VOID\_1} any one individual\newline a pedigreed cat is one whose ancestry is recorded by a cat fancier organization\newline a purebred cat is one whose ancestry contains only individuals of the same breed\newline the cat skull is unusual among \textcolor{red}{VOID\_1} mammals in having very large eye sockets\newline another unusual feature is that the cat cannot produce taurine\newline within groups one cat is usually dominant over the others}